---
title: "Take-Home Exercise 2"
date: "18 February 2023"
date-modified: "18 February 2023"
author: "Tan Wen Yang"
execute: 
  echo: true
  eval: true
  warning: false
editor: visual
---

### Glossary:

Looking at receptiveness

Remove all outer islands

| INDONESIA              | ENGLISH                       |
|------------------------|-------------------------------|
| WILAYAH KOTA           | CITY AREA                     |
| KECAMATAN              | SUBDISTRICT                   |
| KERLURAHAN             | WARD (Local Villages)         |
| SASARAN                | TARGET                        |
| BELUM VAKSIN           | NOT YET VACCINATED            |
| JUMLAH DOSIS           | DOSAGE AMOUNT                 |
| TOTAL VAKSIN DIBERIKAN | TOTAL VACCINE GIVEN           |
| LANSIA DOSIS           | ELDERY DOSAGE                 |
| PELAYAN PUBLIK DOSIS   | PUBLIC SERVANT DOSAGE         |
| GOTONG ROYONG          | KAMPUNG SPIRIT (MUTUAL CO-OP) |
| TENAGA KESEHATAN       | HEALTH WORKERS                |
| TAHAPAN                | STAGES 3 VACCINATION          |
| REMAJA                 | YOUTH VACCINATION             |

### Load the packages in:

| Packages  |                         |
|-----------|-------------------------|
| Lubridate | Handle date/time easily |
|           |                         |
|           |                         |

```{r}
pacman::p_load(sf, tmap, sfdep, tidyverse, readxl, lubridate, r, plyr, ggplot2)
```

# Data-Preprocessing (Geospatial)

```{r}
jakarta_village <- st_read(dsn="data/geospatial", "BATAS_DESA_DESEMBER_2019_DUKCAPIL_DKI_JAKARTA")
```

**Selecting only the first 9 Columns**

Up till JUMLAH_PEN (Total Population)

```{r}
jakarta_village <- jakarta_village[, 0:9]
```

Checking for invalid geometries

```{r}
length(which(st_is_valid(jakarta_village) == FALSE))
```

Check if there are any NA values

```{r}
na_rows <- jakarta_village[apply(is.na(jakarta_village), 1, any),]
na_rows
```

There are empty rows in KESA_KELUR or Kelurahan, which is the target of our analysis. Therefore we will remove it.

```{r}
jakarta_village <- na.omit(jakarta_village,c("DESA_KELUR"))
```

Now we check again

```{r}
na_rows <- jakarta_village[apply(is.na(jakarta_village), 1, any),]
na_rows
```

It appears that there are no more NA rows!

### Check and Transform the CRS of Jakarta

```{r}
st_crs(jakarta_village)
```

**Transform into the appropriate one: DGN95 - ESPG 23845**

```{r}
jakarta_village <- st_transform(jakarta_village, 23845)
```

```{r}
st_crs(jakarta_village)
```

#### Remove Outer Islands

As tasked by the assignment, we have to remove the outer islands. So let's view the map

```{r}
tm_shape(jakarta_village) + 
  tm_polygons("KAB_KOTA")
```

There are a few outer islands in Indonesia, and the ones closest according to Google are:

1.  Kepulauan Seribu Utara (North Thousand Islands)

2.  Kepulauan Seribu Selatan (South Thousand Islands)

We will see if there are any data in the KAB_KOTA (City Regency) that matches both.

```{r}
unique(jakarta_village$"KAB_KOTA")
```

From the above, we can tell that Kepulauan Seribu (Thousand Islands) would encompass both Utara and Selatan (North and South)

#### We will exclude them from our dataset

```{r}
jakarta_village <- filter(jakarta_village, KAB_KOTA != "KEPULAUAN SERIBU")
```

Now we visualize to ensure there are no more outer islands

```{r}
tmap_mode("plot")
tm_shape(jakarta_village) + 
  tm_polygons("KAB_KOTA")
```

It appears they are removed!

# Data-Preprocessing (Aspatial)

We need to process the dataset and only get the data we require. In this case, we need all the city/sub-district/ward areas. To data is cumulative, so we need to subtract from the previous month to get the current vaccinated. When downloaded, it appears that 28 Feb 2022 is missing. So, we'll work with 27 for that month.

To get the number of vaccinated people, we take TARGET - NOT YET VACCINATED column.

We'll take a look at the data we're dealing with by importing the earliest one on July 2021

```{r}
july2021 <- read_xlsx("data/vaccination/Data Vaksinasi Berbasis Kelurahan (31 Juli 2021).xlsx")
```

We glimpse at the data:

```{r}
glimpse(july2021)
```

We only require the first 6 column, and we need to create two more columns. We also remove the first row as that is not necessary

1.  Number vaccinated cumulatively
2.  month

Once we have the number vaccinated cumulatively for that month, we can drop both 'SASARAN' and 'BELUM VAKSIN'.

```{r}
july2021 <- dplyr::select(july2021, 1:6) %>%
  dplyr::slice(-1) %>%
  mutate(total_vaccinated = `SASARAN` - `BELUM VAKSIN`) %>%
  mutate(month = dmy("31-Jul-2021")) %>%
  dplyr::select(-SASARAN, -`BELUM VAKSIN`)
```

Check if our data is correct:

```{r}
july2021
```

Now we will create a function to handle the pre-processing of the data. Mainly translating from Indonesian to English and apply the date into the tibble dataframe.

### Pre-processing Function

The following function will apply the months and date, and also apply the rates of change as defined:

(TARGET - NOT VAX) / TARGET

```{r}

file_list <- list.files(path = "data/vaccination", pattern = "*.xlsx", full.names=TRUE)

process_aspatial_file <- function(file_name) {
  
    #Get the month and year of the file
    pattern <- "(\\b\\d{1,2}\\D{0,3})?\\b(?:Jan(?:uari)?|Feb(?:ruari)?|Mar(?:et)?|Apr(?:il)?|Mei|Jun(?:i)?|Jul(?:i)?|Agu(?:stus)?|Sep(?:tember)?|Okt(?:ober)?|(Nov|Des)(?:ember)?)\\D?(\\d{1,2}\\D?)?\\D?((19[7-9]\\d|20\\d{2})|\\d{2})"

    
    month_dict <- c("Januari"="January","Februari"="February","Maret"="March",
                "April"="April", "Mei"="May", "Juni"="June","Juli"="July",
                "Agustus"="August", "September"="September", "Oktober"="October"                  ,"November"="November","Desember"="December")
    
    matches <- str_extract(file_name, regex(pattern))
    replaced_eng <- str_replace_all(matches, month_dict)
    month_year <- dmy(replaced_eng)
    #month_year_string <- format(month_year, "%d-%m-%Y")
    
    #Calculate the vaccination
    vax_data <- read_xlsx(file_name)
    vax_data <- dplyr::select(vax_data, 1:6) %>%
                dplyr::slice(-1) %>%
                mutate(total_vaccinated_rate = (((`SASARAN` - `BELUM VAKSIN`) / `SASARAN`))) %>%
                mutate(month = month_year) %>%
                dplyr::select(-SASARAN, -`BELUM VAKSIN`)
    
    return(vax_data)
    
  }
```

```{r}
dflist <- lapply(seq_along(file_list), function(x) process_aspatial_file(file_list[x]))
```

#### Get the datatable for the entire Jakarta

```{r}
vacc_jakarta <- ldply(dflist, data.frame)
```

#### Rename the Indonesia names and sort the month column for better understanding:

```{r}
vacc_jakarta <- vacc_jakarta %>% 
  dplyr::rename(
    `village_code`=`KODE.KELURAHAN`,
    `city_area`=`WILAYAH.KOTA`,
    `sub_district`=KECAMATAN, 
    `village`=KELURAHAN
    ) %>%
  arrange(month)
```

#### Check if there are any NA values

```{r}
vacc_jakarta[rowSums(is.na(vacc_jakarta))!=0,]
```

**None! We'll pivot the rows of the dates into columns so it resembles a timeline**

Note that it is a cumulative curve, where the vaccination rates will increase every month.

```{r}
vacc_jakarta <- vacc_jakarta %>% 
                    pivot_wider(names_from=month,
                                values_from=total_vaccinated_rate)
```

We're done with the Aspatial side, now to combine both the geospatial and aspatial data

### Combining Geospatial and Aspatial Data:

From the two datasets, it appears that a good place to perform inner join of the two tables would the the village level. (DESA_KELUR & village)

But first, we need to check to see if they are any discrepancies first before we perform the inner join.

```{r}
jakarta_village_subdis <- c(jakarta_village$DESA_KELUR)
vacc_jakarta_subdis <- c(vacc_jakarta$village)

#Check if they are not in each other
unique(jakarta_village_subdis[!(jakarta_village_subdis %in% vacc_jakarta_subdis)])
```

```{r}
#Check if they are not in each other
unique(vacc_jakarta_subdis[!(vacc_jakarta_subdis %in% jakarta_village_subdis)])
```

There appears to be a mismatch. There are some villages that are NOT present in the vaccine data

#### MISSING VILLAGES IN VACCINE DATA:

| Missing Villages  |
|-------------------|
| PULAU KELAPA      |
| PULAU HARAPAN     |
| PULAU PARI        |
| PULAU TIDUNG      |
| PULAU UNTUNG JAWA |
| PULAU PANGGANG    |

These are villages from the outer islands, therefore should be excluded, but inner join operation will automatically drop and exclude them. So we do not need to worry about them!

We need to change and modify the following: (From quick research, it appears there are some spelling errors on the aspatial data side)

| Geospatial           | Aspatial                      |
|----------------------|-------------------------------|
| BALEKAMBANG          | BALE KAMBANG                  |
| HALIM PERDANA KUSUMA | HALIM PERDANA KUSUMAH (error) |
| JATIPULO             | JATI PULO                     |
| KRENDANG             | KERENDANG (ERROR)             |
| KRAMATJATI           | KRAMAT JATI                   |
| PALMERIAM            | PAL MERIAM                    |
| PINANGRANTI          | PINANG RANTI                  |
| TENGAH               | KAMPUNG TENGAH                |
| RAWAJATI             | RAWA JATI                     |

### Make the modifications:

We will change the vacc_jakarta set as there are misspellings (Kerendang should be Krendang) in the vacc_jakarta set to match with the geospatial set.

```{r}
vacc_jakarta$village[vacc_jakarta$village == 'BALE KAMBANG'] <- 'BALEKAMBANG'
vacc_jakarta$village[vacc_jakarta$village == 'HALIM PERDANA KUSUMAH'] <- 'HALIM PERDANA KUSUMA'
vacc_jakarta$village[vacc_jakarta$village == 'JATI PULO'] <- 'JATIPULO'
vacc_jakarta$village[vacc_jakarta$village == 'KERENDANG'] <- 'KRENDANG'
vacc_jakarta$village[vacc_jakarta$village == 'KRAMAT JATI'] <- 'KRAMATJATI'
vacc_jakarta$village[vacc_jakarta$village == 'PAL MERIAM'] <- 'PALMERIAM'
vacc_jakarta$village[vacc_jakarta$village == 'PINANG RANTI'] <- 'PINANGRANTI'
vacc_jakarta$village[vacc_jakarta$village == 'KAMPUNG TENGAH'] <- 'TENGAH'
vacc_jakarta$village[vacc_jakarta$village == 'RAWA JATI'] <- 'RAWAJATI'
```

Check again to make sure modification are done correctly:

```{r}
jakarta_village_subdis <- c(jakarta_village$DESA_KELUR)
vacc_jakarta_subdis <- c(vacc_jakarta$village)

unique(jakarta_village_subdis[!(jakarta_village_subdis %in% vacc_jakarta_subdis)])
```

```{r}
unique(vacc_jakarta_subdis[!(vacc_jakarta_subdis %in% jakarta_village_subdis)])
```

All good, now we can perform inner join.

```{r}
combined_jakarta <- left_join(jakarta_village, vacc_jakarta,
                              by=c("DESA_KELUR"="village"))

combined_jakarta <- combined_jakarta %>% 
  dplyr::select(1:9, 13:24)
```

#### Visualize the Monthly Rate of Vaccination:

```{r}
trace_val <- c("2021-07-31", "2021-08-31", "2021-09-30", "2021-10-31", "2021-11-30", "2021-12-31", "2022-01-31", "2022-02-27", "2022-03-31", "2022-04-30", "2022-05-31", "2022-06-30")

plot_creation <- function(df, month_name){
  tmap_mode("plot")
  tm_shape(df) +
    tm_borders(alpha=0.5)+
    tm_fill(month_name, 
            title="Vaccination Rate",
            breaks=c(0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9))+
  tm_layout(main.title = month_name,
          main.title.position = "center",
          main.title.size = 1.2,
          legend.height = 0.45, 
          legend.width = 0.35,
          frame = TRUE)
}
```

#### View the map interactively (June 2022)

```{r}
tmap_mode("view")+
tm_shape(combined_jakarta) +
    tm_borders(alpha=0.5)+
    tm_polygons("2022-06-30",
                id="DESA_KELUR", 
                breaks=c(0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9),
                title="Vaccination Rate")

```

```{r}
tmap_arrange(plot_creation(combined_jakarta, "2021-07-31"),
             plot_creation(combined_jakarta, "2021-08-31"),
             plot_creation(combined_jakarta, "2021-09-30"),
             plot_creation(combined_jakarta, "2021-10-31"))
```

```{r}
tmap_arrange(plot_creation(combined_jakarta, "2021-11-30"),
             plot_creation(combined_jakarta, "2021-12-31"),
             plot_creation(combined_jakarta, "2022-01-31"),
             plot_creation(combined_jakarta, "2022-02-27"))
```

```{r}
tmap_arrange(plot_creation(combined_jakarta, "2022-03-31"),
             plot_creation(combined_jakarta, "2022-04-30"),
             plot_creation(combined_jakarta, "2022-05-31"),
             plot_creation(combined_jakarta, "2022-06-30"))
```

## Describe the spatial patterns revealed by the maps:

From a glance, it appears that during the first three months, during 2021, only a few city states has a high rate of vaccination. However, in subsequent months, the neighboring village sub-district all begin to catch up save for a few. Eventually ending in 2022 June with most of the people vaccinated in Jakarta.

The village closest to the villages with higher vaccination rates are the second to start receiving the vaccination doses, as they show demonstrably higher vaccination rates compared to other villages (October 2021). They are then slowly doled out across nearby villages until most villages in Jakarta are almost fully vaccinated.

Interestingly, during this period of measure (July), there was an exponential spike [(Reuters)](https://www.reuters.com/world/asia-pacific/indonesia-emergency-covid-19-curbs-be-effective-july-2-20-minister-2021-07-01/) in Covid-19 cases with heavy bans and movement restrictions across the board with recorded 2.2 million cases. This might explain why the almost equal distribution in terms of vaccination rates across the board in Jakarta. The only two villages, seemingly with lower rates compared to the rest of the city are:

1.  Senayan

2.  Kebon Melati

3.  Petamburan

### Plot the rate of change using tmap

```{r}
#trace_val <- c("2021-07-31", "2021-08-31", "2021-09-30", "2021-10-31", "2021-11-30", "2021-12-31", "2022-01-31", "2022-02-27", "2022-03-31", "2022-04-30", "2022-05-31", "2022-06-30")

#jakarta_combined_long <- combined_jakarta %>% 
#  pivot_longer(cols=trace_val, names_to="Month_Vac", values_to = #"total_vaccinated_rate")
```

# Local Gi\* Analysis

Over here we will compute the local Gi\* values for the monthly vaccination rates. Gi\* only works for distance-based matrix and queen contiguity. This is to detect hot/cold spots for villages with high/low covid-19 vaccinations respectively. We will do it in a time-series to understand how the vaccinations are doled out across the country overtime.

**Intepreting Gi\* Values:**

1.  Gi\* \> 0: Grouping of areas with values higher than average
2.  Gi\* \< 0: Grouping of areas with values lower than average

The values represents the intensity of grouping.

**Set the seed for reproducability**

```{r}
set.seed(1234)
```
