---
title: "Hands-on Exercise 4"
date: "28 January 2023"
format: html
date-modified: "`r Sys.Date()`"
number-sections: true

execute:
  message: true
  warning: false
  echo: true
  eval: true
  
editor: visual
---

Loading packages

```{r}
pacman::p_load(maptools, sf, raster, spatstat, tmap)
```

Load Spatial Data

```{r}
childcare_sf <- st_read("data/childcare.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
sg_sf <- st_read(dsn = "data", layer="CostalOutline")
```

```{r}
mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")
```

### Mapping Geospatial Set

```{r}
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()

```

```{r}
tmap_mode('plot')
```

### Convert to spatial class

```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

```{r}
childcare
```

```{r}
mpsz
```

```{r}
sg

```

### Converting the Spatial\* class into generic sp format

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

```{r}
childcare_sp
```

```{r}
sg_sp
```

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```

```{r}
plot(childcare_ppp)
```

```{r}
summary(childcare_ppp)
```

Check for duplication

```{r}
any(duplicated(childcare_ppp))
```

Number of Coincidence Points

```{r}
multiplicity(childcare_ppp)
```

### Number of locations with more than one point event

```{r}
sum(multiplicity(childcare_ppp) > 1)

```

### Viewing them

```{r}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)

```

### Jittering Approach to identify duplicates 

Ensure they don't occupy the same space

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

```{r}
any(duplicated(childcare_ppp_jit))
```

### Creating ***owin*** object

Used to confine and represent a region using polygons

```{r}
sg_owin <- as(sg_sp, "owin")
```

```{r}
plot(sg_owin)

```

Summary stat

```{r}
summary(sg_owin)

```

### Combining point events object and owin object

Extract childcare events that are located within Singapore by using the code chunk below:

```{r}
childcareSG_ppp = childcare_ppp[sg_owin]
```

```{r}
summary(childcareSG_ppp)
```

```{r}
plot(childcareSG_ppp)
```

### First-order Spatial Point Patterns Analysis

#### KDE (Kernel Density Estimation)

```{r}
kde_childcareSG_bw <- density(childcareSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 

```

```{r}
plot(kde_childcareSG_bw)

```

Finding the bandwidth to compute the kde layer

```{r}
bw <- bw.diggle(childcareSG_ppp)
bw

```

#### Rescaling the KDE values

```{r}
childcareSG_ppp.km <- rescale(childcareSG_ppp, 1000, "km")
```

```{r}
kde_childcareSG.bw <- density(childcareSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG.bw) 
```

#### Working with different automatic badwidth methods

```{r}
 bw.CvL(childcareSG_ppp.km)
```

```{r}
bw.scott(childcareSG_ppp.km)
```

```{r}
bw.ppl(childcareSG_ppp.km)
```

```{r}
bw.diggle(childcareSG_ppp.km)
```

Compare output between diggle and ppl methods

```{r}
kde_childcareSG.ppl <- density(childcareSG_ppp.km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```

### Different kernel methods 

Gaussian, Epanechnikov, Quartic and Dics

```{r}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```

### Fixed and Adaptive KDE

#### Computing KDE using fixed bandwidth

```{r}
kde_childcareSG_600 <- density(childcareSG_ppp.km, sigma=0.6, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG_600)
```

#### Computing KDE by using adaptive bandwidth

```{r}
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
plot(kde_childcareSG_adaptive)
```

#### Compare fixed and adaptive

```{r}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")

```

#### Converting KDE output into Grid Object

```{r}
gridded_kde_childcareSG_bw <- as.SpatialGridDataFrame.im(kde_childcareSG.bw)
spplot(gridded_kde_childcareSG_bw)
```

#### Converting gridded output into raster

```{r}
kde_childcareSG_bw_raster <- raster(gridded_kde_childcareSG_bw)
```

NA crs (projection system)

```{r}
kde_childcareSG_bw_raster
```

Assign Projection systems

```{r}
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster

```

### Visualising output in tmap

```{r}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)

```

#### Comparing Spatial Point Patterns using KDE

Extracting study area

```{r}
pg = mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
tm = mpsz[mpsz@data$PLN_AREA_N == "TAMPINES",]
ck = mpsz[mpsz@data$PLN_AREA_N == "CHOA CHU KANG",]
jw = mpsz[mpsz@data$PLN_AREA_N == "JURONG WEST",]

```

Plot target planning areas

```{r}
par(mfrow=c(2,2))
plot(pg, main = "Punggol")
plot(tm, main = "Tampines")
plot(ck, main = "Choa Chu Kang")
plot(jw, main = "Jurong West")
```

### Convert spatial point data frame into generic sp format

```{r}
pg_sp = as(pg, "SpatialPolygons")
tm_sp = as(tm, "SpatialPolygons")
ck_sp = as(ck, "SpatialPolygons")
jw_sp = as(jw, "SpatialPolygons")
```

#### Create owin object

```{r}
pg_owin = as(pg_sp, "owin")
tm_owin = as(tm_sp, "owin")
ck_owin = as(ck_sp, "owin")
jw_owin = as(jw_sp, "owin")
```

Combine childcare points and study area

```{r}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
```

Rescale to transform unit of measurement from metre to kilometer

```{r}
childcare_pg_ppp.km = rescale(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale(childcare_jw_ppp, 1000, "km")
```

Plot the study areas

```{r}
par(mfrow=c(2,2))
plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```

### Computing KDE

```{r}
par(mfrow=c(2,2))
plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")

```

### Computing fixed bandwidth kde

```{r}
par(mfrow=c(2,2))
plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chou Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Jurong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
```

## Nearest Neighbour Analysis

Testing spatial point patterns using Clark and Evans

Across the board in sg: Good spread

```{r}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

### Clark and Evans Test: Choa chu kang planning area

Result: VERY clustered

```{r}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

### Clark and Evans Test: Tampines planning area

Somewhat clustered

```{r}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```
